Efficient Computation of Extended Sources

### 2022-03-28 repro/papers/GEM22/newstuff.tex
first commit

%The energy source $\bff$ may also be largely undetermined, apart from some known characteristics such as localization in space and/or time. In fact, additional source degrees of freedom, beyond those needed to describe physically realized sources, may be useful in rendering the FWI problem \ref{eqn:fwi} more amenable to numerical solution, via so-called extended modeling (see \cite{geoprosp:2008}, \cite{LeeuwenHerrmannWRI:13}, \cite{HuangNammourSymesDollizal:SEG19}, and many references cited there). Therefore it is natural to view $\bff$ as also an unknown in formulating the problem \ref{eqn:fwi} via nonlinear least squares:
A simple nonlinear least squares formulation is:
\begin{equation}
\label{eqn:ols}
\mbox{choose } {\bf c} \mbox{ to minimize } \|PL[{\bf c}]^{-1}\bff -d \|^2.
\end{equation}
Practical optimization formulations typically augment the objective in
\ref{eqn:ols} by additive penalties or other constraints.

@@

\cite{HuangNammourSymesDollizal:SEG19} present an overview of the
literature on source extension methods, describing a variety of
methods to add degrees of freedom to physical source model. The present paper
concerns {\em surface source extension}: physical sources are
presumed to be concentrated at points $\bx_s$ in space, whereas their extended
counterparts are permitted to spread energy over surfaces containing
the physical source locations. A simple choice for the penalty
operator $A$ is then multiplication by the distance $|\bx-\bx_s|$ to the physical
source location:
\begin{equation}
  \label{eqn:penop}
  (A\bff)(\bx,t) = |\bx-\bx_s|\bff(\bx,t)
\end{equation}
I shall use this choice of penalty operator whenever a specific choice
is necessary in the development of the theory below.

This paper presents a numerically efficient approach to solving the
{\em source subproblem} of problem \ref{eqn:esi}:
\begin{equation}
\label{eqn:esis}
\mbox{given } {\bf c}, \mbox{ choose } \bff \mbox{ to minimize }
\|PL[{\bf c}]^{-1}\bff -d \|^2 + \alpha \|A\bff\|^2 
\end{equation}
Solution of this subproblem is an essential component of {\em variable
  projection} algorithms for solution of the nonlinear inverse problem
\ref{eqn:esi}. Variable projection is not merely a convenient choice
of algorithm for this purpose: it is in some sense essential, see for
example \cite{Symes:SEG20}. It replaces the nonlinear
least squares problem \ref{eqn:esi} with a {\em reduced} problem, to
be solved iteratively. Each iteration involves solution of the
subproblem \ref{eqn:esis}. Therefore efficient solution of the
subproblem is essential to efficient solution of the nonlinear problem
via variable projection.

The penalty operator $A$ defined in \ref{eqn:penop} is linear, so the source
subproblem is a linear least squares problem. Under some additional
assumptions to be described below, I shall show how to construct an
accurate approximate solution operator for problem
\ref{eqn:esis}. This approximate solution operator may be used to
accelerate Krylov space methods for the solution of the source
subproblem \ref{eqn:esis}. Numerical examples suggest the
effectiveness of this acceleration.

@@

The obvious computation of the pressure-to-source map
- prescribe the pressure, solve the wave equation with this boundary
condition, read off the equivalent source - suffers from intrinsic
numerical inaccuracy. I suggest an alternative computationally
feasible approach.

@@

Most of what follows is valid for any space dimension $n >0$. The
coefficient vector $\bf{c}=(\kappa,\rho)$ is defined throughout space
$\bR^n$, the state vector $\bu$ throughout space-time
$\bR^{n+1}$. Whenever convenient for mathematical manipulations,
$n=3$: for instance, I will write $\bx=(x,y,z)^T$ for the spatial
coordinate vector, and refer to the third (vertical) coordinate of
particle velocity as $v_z$. Examples
later in this paper will use $n=2$ for computational convenience.

@@

The surface source extension replaces point sources on or near a
surface in $\bR^3$ with source functions confined to the same
surface. The simplest example of this extended geometry specifies a
plane $\{(x,y,z,t): z=z_s\}$ at source depth $z_s$ as the surface. For
acoustic modeling, surface sources are combinations of constitutive law
defects and loads normal to the surface, localized on $z=z_s$. That
is, right-hand sides in the system $L\bu=\bff$ take the form
$\bff(\bx,t) = (h_s(x,y,t)\delta(z-z_s),
f_s(x,y,t)\bf{e}_z\delta(z-z_s))^T$ for scalar defect $h_s$ and normal
force $f_s$ ($\bf{e}_z=(0,0,1)$). With the choice $L$ given in
\ref{eqn:awemat}, the causal/anti-causal wave system $L\bu^{\pm}=\bff$
takes the form
\begin{eqnarray}
\label{eqn:awepm}
\frac{1}{\kappa}\frac{\partial p^{\pm}}{\partial t} & = & - \nabla \cdot \bv^{\pm} +
h_s \delta(z-z_s), \nonumber \\
\rho\frac{\partial \bv^{\pm}}{\partial t} & = & - \nabla p^{\pm} +
                                                f_s{\bf e} \delta(z-z_s),\nonumber \\
p^{\pm} & =& 0 \mbox{ for } \pm t \ll 0,\nonumber\\ 
\bv^{\pm} & = & 0 \mbox{ for } \pm t \ll 0.
\end{eqnarray}

\noindent {\bf Remark:} In system \ref{eqn:awepm} and many similar
systems to follow, I will use the shorthand
\[
  p^+ = 0 \mbox{ for } t \ll 0 
\]
to mean that $p^+$ is {\em causal}, that is,
\[
  \mbox{For some } T \in \bR, p^+(\cdot,t) = 0 \mbox{ for all } t <
  T.
\]
Similarly,
\[
  p^- = 0 \mbox{ for } t \gg 0 
\]
signifies that $p^-$ is anti-causal.

@@

\section{2D Examples}

\inputdir{project}

To illustrate the structure described in the preceding section, I
introduce two 2D acoustic models, one spatially homogeneous, the other
with highly refractive. The first, homogenous model has $\kappa = 4$
GPa and $\rho = 1$ g/cm$^3$ throughout a rectangular domain of size 8 km ($x$) $\times$ 4 km
($z$). The second, refractive, model is a perturbation of the first by
a low-velocity acoustic lens positioned in the center of the
rectangle. To produce this structure, the density is chosen
homogeneous as in the first model, while the bulk modulus decreases to
from 4 GPa outside the lens to 1.6 GPA in its center, as shown in Figure \ref{fig:bml0}.

Discretization is conventional, with a rectangular grid and staggered
finite difference scheme \cite[]{Vir:84} of order 2 in time and
2$k$ in space; for most of the experiments reported below, $k=4$.
Sampling operators such as $P_r$ are implemented via linear
interpolation, and source insertion via adjoint linear interpolation
(as noted above, in the continuum limit, sources are represented via
adjoint sampling). Steps in $x$ and $z$ are the same. Most results use
$\Delta x = 20$ m, and limit the temporal frequency of the computed
traces to 12.5 Hz. We also explore the dependence of a few results on frequency, using
$\Delta x = 10$ m and $5$ m, to accomodate 25 and 50 Hz respectively,
maintaining 8 samples per wavelength. All computations are carried out
in single precision.

\cite{GeoPros:11} gives a description of the code
implementation, out-of-date in a few respects but overall accurate.

The horizontal line of receivers sits at depth $z_r = $ 1000 m, as
shown in Figure \ref{fig:bml0}. Receiver $x$ ranges from $2000$ to
$6000$ m. A single point (physical) source appears in these
experiments, located at $x_s=3500$ m, $z_s=3000$ m, as also indicated
in the figure. The source time function is a bandpass filter with
corner frequencies $1, 2.5, 7.5, 12.5$ Hz for the lowest frequency
source, and scaled as appropriate for examples with higher frequency
and finer sampling.  Extended sources are confined to the horizontal
line through the physical source position, that is $z_s = 3000$ m,
over a 4 km interval starting at $x_r=$ 2000 m. Note that we have
reversed the order relation between $z_s$ and $z_r$ described in the
text ($z_s<z_r$). This difference is immaterial.

The point source pressure data generated by this configuration for the
homogeneous model is displayed in Figure \ref{fig:recphh0}, for the
lens model in \ref{fig:recplh0}. Triplication of arrivals is evident in the latter.

While inversion of pressure data is the main object of this exercise,
normal velocity ($v_z$) data will play an important role, so I display
the corresponding gathers in Figures \ref{fig:recvzhh0} and \ref{fig:recvzlh0}.

\section{Adjoints}

It follows from the adjoint state method (see Appendix A for details) that
\begin{equation}
  \label{eqn:sadj1}
  ({\cal S}^{\pm}_{z_s, z_r})^T = -{\cal S}^{\mp}_{z_r,z_s}
\end{equation}

Define $R$ to be the {\em time-reversal operator} on functions of
space-time, $Rf(\bx,t) = f(\bx,-t)$, and ${\cal R}$ to be the {\em
  acoustic field time-reversal operator}
\begin{equation}
  \label{eqn:trdef}
  {\cal R} \left(
    \begin{array}{c}
      p\\
      \bv
    \end{array}
  \right) =
  \left(
    \begin{array}{c}
      Rp\\
      -R\bv
    \end{array}
  \right)
\end{equation}
Then 
\begin{equation}
  \label{eqn:trsadj}
  {\cal R}{\cal S}^{\mp} = -{\cal S}^{\pm}_{z_r,z_s}{\cal R}
\end{equation}
Since $R^2 = I$ and ${\cal R}^2 = I$, the identities \ref{eqn:sadj} and \ref{eqn:trsadj} imply that

\begin{equation} 
  \label{eqn:trtr}
 ({\cal S}^{\pm}_{z_s,z_r})^T = {\cal R}{\cal S}_{z_r,z_s}^{\pm}{\cal R}=
 -{\cal S}^{\mp}_{z_r,z_s}.
\end{equation}

The relation \ref{eqn:trtr} implies that
\begin{eqnarray}
  (S^{\pm}_{z_s,z_r})^T &=& -S^{\mp}_{z_r,z_s} \nonumber\\
                        &=& R S^{\pm}_{z_r,z_s}R, \nonumber\\
    (V^{\pm}_{z_s,z_r})^T &=& -V^{\mp}_{z_r,z_s} \nonumber\\
                        &=& R V^{\pm}_{z_r,z_s}R.
                            \label{eqn:trtrcomp}
\end{eqnarray}

>>>>

A simple nonlinear least squares formulation, including estimation of the source, is:
\begin{equation}
\label{eqn:ols}
\mbox{choose } {\bf c}, \bff \mbox{ to minimize } \|PL[{\bf c}]^{-1}\bff -d \|^2.
\end{equation}
Practical optimization formulations typically augment the objective in
\ref{eqn:ols} by additive penalties or other constraints.

@@

\cite{HuangNammourSymesDollizal:SEG19} present an overview of the
literature on source extension, describing a variety of
methods to add degrees of freedom to physical source model. The present paper
concerns {\em surface source extension}: physical sources are
presumed to be concentrated at points $\bx_s$ in space, whereas their extended
counterparts are permitted to spread energy over surfaces containing
the physical source locations. A simple choice for the penalty
operator $A$ is then multiplication by the distance $|\bx-\bx_s|$ to the physical
source location:
\begin{equation}
  \label{eqn:penop}
  (A\bff)(\bx,t) = |\bx-\bx_s|\bff(\bx,t)
\end{equation}
I shall use this choice of penalty operator whenever a specific choice
is necessary in the development of the theory below.

This paper presents a numerically efficient approach to solving the
{\em source subproblem} of problem \ref{eqn:esi}:
\begin{equation}
\label{eqn:esis}
\mbox{given } {\bf c}, \mbox{ choose } \bff \mbox{ to minimize }
\|PL[{\bf c}]^{-1}\bff -d \|^2 + \alpha \|A\bff\|^2 
\end{equation}
Solution of this subproblem is an essential component of {\em variable
  projection} algorithms for solution of the nonlinear inverse problem
\ref{eqn:esi}. Variable projection is not merely a convenient choice
of algorithm for this purpose: it is in some sense essential, see for
example \cite{Symes:SEG20}. It replaces the nonlinear
least squares problem \ref{eqn:esi} with a {\em reduced} problem, namely the function of ${\bf c}$ defined by the minimum of the subproblem objective \ref{eqn:esis}. Evaluation of this reduced objective requires solution of the
subproblem \ref{eqn:esis}. Therefore efficient solution of this
subproblem is essential to efficient solution of the nonlinear problem
via variable projection.

The penalty operator $A$ defined in \ref{eqn:penop} is linear, so the source
subproblem is a linear least squares problem. Under some additional
assumptions to be described below, I shall show how to construct an
accurate approximate solution operator for problem
\ref{eqn:esis}. This approximate solution operator may be used to
accelerate (``precondition'') Krylov space methods for the solution of the source
subproblem \ref{eqn:esis}. Numerical examples suggest the
effectiveness of this acceleration.

Beyond accelerated Krylov iteration for solution of problem \ref{eqn:esis}, the structure of the preconditioner is key to understanding the relation of the extended problem \ref{eqn:esi} to traveltime tomography, which in turn explains its ability to avoid cycle-skipping \cite[]{Symes:22b}. The same structure enables efficient and accurate evaluation of the reduced objective gradient, another essential task in the computation solution of the extended inverse problem \cite[]{Symes:22c}.

@@

%The obvious computation of the pressure-to-source map
%- prescribe the pressure, solve the wave equation with this boundary
%condition, read off the equivalent source - suffers from intrinsic
%numerical inaccuracy. I suggest an alternative computationally
%feasible approach.

@@

Most of what follows is valid for any space dimension $n >0$. The
coefficient vector $\bf{c}=(\kappa,\rho)$ is defined throughout space
$\bR^n$, and are assumed to be smooth, in fact $\log \kappa, \log \rho \in C^{\infty}(\bR^d)$ and uniformly bounded. The same is true of the wave velocity $c = \sqrt{\kappa/\rho}$. The state vector $\bu$ is a vector-valued distribution on space-time
$\bR^{n+1}$: $\bu = (p,\bv)$, in which the pressure field $p$ is scalar and the (particle) velocity field $\bv$ is an $n$-vector field. Whenever convenient for mathematical manipulations,
$n=3$: for instance, I will write $\bx=(x,y,z)^T$ for the spatial
coordinate vector, or $(\bx,t) = (x,y,z,t)$ for the space-time coordinate vector, and refer to the third (vertical) coordinate of
particle velocity as $v_z$. Examples
later in this paper will use $n=2$ for computational convenience.

@@

\noindent {\bf Remark:} I will use the shorthand
\[
  \bu  = 0 \mbox{ for } t \ll 0 
\]
to mean that $\bu$ is {\em causal}, that is,
\[
  \mbox{For some } T \in \bR, \bu(\cdot,t) = 0 \mbox{ for all } t <
  T.
\]
Similarly,
\[
  \bu = 0 \mbox{ for } t \gg 0 
\]
signifies that $\bu$ is anti-causal.

It has been understood for decades that initial/boundary value problems for the system $L\bu=\bff$ have well-behaved solutions. In particular, \cite{BlazekStolkSymes:13} state results which imply the following

\begin{theorem}
  \label{thm:basicp}
  Denote by $\Omega \subset \bR^n$ a bounded domain with smoothly embedded oriented boundary $\partial \Omega$, and $V=H^1_0(\Omega) \times H^1_{\rm div}(\Omega)$. 
  \begin{itemize}
  \item[i. ] Suppose that $\bff = (f,{\bf g}) \in (L^2_{\rm comp}(\Omega \times {\bf R}))^{n+1}$. Then the system $L\bu=\bff$ has a unique causal weak solution $\bu = (p,\bv) \in (L_{\rm loc}^2(\Omega \times \bR))^{n+1}$: for any ${\bf \phi} = (\phi, {\bf \psi}) \in (C_0^{\infty}(V))^{n+1}$,
    \begin{equation}
      \label{eqn:weak}
      \int \left(\frac{1}{\kappa} p \frac{\partial \phi}{\partial t} + \rho \bv \cdot \frac{\partial{\bf \psi}}{\partial t} + \bv \cdot \nabla \phi + p \nabla \cdot {\bf \psi} \right) = \int (f\phi + {\bf g} \cdot {\bf \psi}).
    \end{equation}
  \item[ii. ] If in addition $1 \le k \in {\bf N}$ and
    $\partial^k_t\bff \in (L^2(\Omega \times \bR))^{n+1}$, then $\bu
    \in C^k(\bR, (L^2(\Omega))^{n+1}) \cap C^{k-1}(\bR,V)$: that is,
    $\bu$ is a strong solution, satisfying $L\bu=\bff$ almost everywhere..
  \item[iii. ] 
    Suppose that $R>0$ and $\mbox{supp }\bff \subset [T_0, T_1] \times B_R(\bx_0)$. Then $\bu(\cdot,t)=0$ for $t<T_0$, and for each $t \in [T_0,\infty)$, $\mbox{supp }\bu \subset B_R(\bx_0) + B_{c_{\rm max}(t-T_0)}({\bf 0})$.
  \end{itemize}
\end{theorem}

\begin{cor}
  \label{thm:weakcor}
  Let $\alpha_{\epsilon}$ be a Dirac family in $C_0^{\infty}(\bR)$,
  and $\bu$ be a causal weak solution of $L\bu=\bff \in (L^2_{\rm
    comp}(\Omega \times \bR))^{n+1}$. Then for all $\epsilon > 0$,
  $\alpha_{\epsilon} *_t \bu$ is a strong solution with well-defined and vanishing restriction
  to $\partial \Omega \times \bR$.
\end{cor}

Denote by ${\bf \nu}(\bx)$ the outward unit normal (co)vector at $ \bx
\in \partial \Omega$.

\begin{theorem}
  \label{thm:basicv}
  Let  $H^1_{{\rm div},0}(\Omega)$ denote the subspace of $H^1_{\rm
    div}(\Omega)$ consisting of vector fields $\bv$ satisfying $\bv
  \cdot {\bf \nu} = 0$ on $\partial \Omega$. Define
  $W = H^1(\Omega) \times H^1_{{\rm div},0}(\Omega)$.
Then the same conclusions as listed in Theorem \ref{thm:basicp} hold,
with $V$ replaced by $W$.
\end{theorem}
\begin{proof}
  This result follows directly from Theorem 1 in
  \cite[]{BlazekStolkSymes:13}, once it is shown that the spatial part
  of $L$ is skew-adjoint on $W$. The analogous statement for
  $V$ is proven in \cite[]{BlazekStolkSymes:13}, Appendix A, and a
  similar proof establishes the same property of $W$.
\end{proof}

\begin{remark} The cited reference \cite[]{BlazekStolkSymes:13} requires only that the coefficients ${\bf c}$ be bounded and measureable. The arguments developed here, however, depend on approximation of solutions by means of geometric optics, hence the additional requirement of coefficient smoothness.
\end{remark}

The surface source extension replaces point sources on or near a
surface in $\bR^3$ with distributions supported on the same
surface. The simplest example of this extended geometry specifies a
plane $\{(x,y,z,t): z=z_s\}$ at source depth $z_s$ as the surface. For
acoustic modeling, surface sources are combinations of constitutive law
defects and loads normal to the surface, localized on $z=z_s$. That
is, right-hand sides in the system $L\bu=\bff$ take the form
$\bff(\bx,t) = (h_s(x,y,t)\delta(z-z_s),
f_s(x,y,t)\bf{e}_z\delta(z-z_s))^T$ for scalar defect $h_s$ and normal
force $f_s$ ($\bf{e}_z=(0,0,1)$). With the choice $L$ given in
\ref{eqn:awemat}, the causal/anti-causal wave system $L\bu^{\pm}=\bff$
takes the form
\begin{eqnarray}
\label{eqn:awepm}
\frac{1}{\kappa}\frac{\partial p^{\pm}}{\partial t} & = & - \nabla \cdot \bv^{\pm} +
h_s \delta(z-z_s), \nonumber \\
\rho\frac{\partial \bv^{\pm}}{\partial t} & = & - \nabla p^{\pm} +
                                                f_s{\bf e}_z \delta(z-z_s),\nonumber \\
p^{\pm} & =& 0 \mbox{ for } \pm t \ll 0,\nonumber\\ 
\bv^{\pm} & = & 0 \mbox{ for } \pm t \ll 0.
\end{eqnarray}

The singular right-hand sides appearing in the systems \ref{eqn:awepm} do not conform to the conditions of the basic well-posedness statement \ref{thm:basicp}, so the first task is to formulate conditions under which well-behaved solutions exist. This is most conveniently done by first developing conditions under which an inhomogeneous boundary value problem has well-behaved solutions. The pressure field on the surface $\partial \Omega \times \bR$ uniquely determines the finite energy ($L^2$) causal acoustic field in the interior, as follows from simple energy estimates, but cannot be specified arbitrarily. A finite energy solution exists if the pressure field on the boundary is {\em incoming} (corresponding to {\em downgoing} in seismic parlance), in the sense that its Fourier components decrease rapidly with increasing frequency except at initial conditions for rays entering $\Omega$ at a positive angle, with distance to the boundary increasing with increasing time. A precise statement depends on a parameter $\delta >0$, essentially a bound on the direction cosine relative to the tangent plane of the boundary.

\begin{definition}
  \label{thm:definc}
  Given $\delta >0$, $d \in L^2_{\rm comp}(\partial \Omega \times \bR)$ is $\delta-${\em incoming} if there exists a conic subset $\Gamma \subset T^*(\partial \Omega \times \bR)$ and $H \in OPS^0(\partial \Omega \times \bR)$ satsifying the conditions:
  \begin{itemize}
  \item[1. ] $H$ is elliptic in $\Gamma$;
  \item[2. ] for every $(\bx,t,\xi,\omega) \in \Gamma$, there is a
    unique $z \in \bR$ and $\psi(\bx,{\bf \xi},\omega) = \frac{{\bf \xi}}{\omega} + z {\bf \nu}(\bx)  \in
    \bR^n$ so that $(\bx,t,\psi(\bx,{\bf \xi},\omega),\omega)$ is
    characteristic for $L$, $|z| > \delta$, and $z\omega < 0$; and
  \item[3. ] $(I-H)d \in H^2(\partial \Omega \times \bR)$.
  \end{itemize}
\end{definition}

For $\epsilon > 0$, denote by $\Omega_{\epsilon}$ the collar neigborhood of $\partial \Omega$ defined by
\begin{equation}
  \label{eqn:collar}
  \Omega_{\epsilon} = \{\bx + z {\bf \nu}(\bx): 0 \ge z \ge  -\epsilon\}.
\end{equation}
For $\epsilon$ sufficiently small, $\Omega_{\epsilon}$ is a smoothly bounded submanifold of $\Omega$.

\begin{theorem}
  \label{thm:dirprob}
  Suppose that $\delta>0$ and $d \in L^2_{\rm comp}(\partial \Omega
  \times \bR)$ is $\delta-${\em incoming}. Then there exists a unique
  causal weak solution $\bu = (p,\bv) \in (L^2_{\rm loc}(\Omega \times
  \bR)^{n+1}$ of $L\bu=0$ in $\Omega$, for which $p$ assumes the boundary value $d$, in the sense that for every $\phi \in C_0^{\infty}(\bR^{n+1})$,
  \begin{equation}
    \label{eqn:weakbdry}
    \lim_{\epsilon \rightarrow 0}(\alpha_{\epsilon} *_t p)|_{\partial
      \Omega \times \bR}  = d.
  \end{equation}
\end{theorem}

\begin{proof}
  Choosing $\epsilon$ sufficiently small, construct an extension $d_1\in H^2(\Omega_{\epsilon}\times \bR)$ of $(I-H)d$, and cut it off smoothly to produce an extension, also denoted $d_1$, to all of $\Omega \times \bR$. Define $(p_1,\bv_1)$ to be the weak solution of
\begin{eqnarray}
\label{eqn:awepm1}
\frac{1}{\kappa}\frac{\partial p_1}{\partial t} & = & - \nabla \cdot \bv_1 -
\frac{1}{\kappa}\frac{\partial d_1}{\partial t}, \nonumber \\
\rho\frac{\partial \bv_1}{\partial t} & = & - \nabla p_1 \\
p_1& =& 0 \mbox{ for } t \ll 0,\nonumber\\ 
\bv_1 & = & 0 \mbox{ for } t \ll 0.
\end{eqnarray}
From Theorem \ref{thm:basicp} with $k=1$, conclude that $(p_1,\bv_1)
\in H^1_0(\Omega \times \bR) \times H^1_{\rm div}(\Omega \times \bR)$
is a strong solution and in particular that the trace of
$p_1|_{\partial \Omega \times \bR} \in L^2_{\rm loc}(\partial \Omega \times \bR)$ is
well-defined, and in fact
\begin{equation}
  \label{eqn:smparttrace}
  p_1|_{\partial \Omega \times \bR} = (I-H)d.
\end{equation}

Let $\{({\cal U}_i, \chi_i): i=1,...,N\}$ be a coordinate atlas for
the oriented submanifold $\partial \Omega \subset \bR^n$. That is,
$\{{\cal U}_i: i=1,...N\}$ is a locally finite open cover of $\partial \Omega$, 
$\chi_i:{\cal U}_i \rightarrow \bR^{n-1}$ is a diffeomorphism for each $i$, and
$\partial \Omega \cap {\cal U}_i = \chi_i^{-1}\{\bx \in \bR^n:
x_n=0\}$. Define $\tilde{\chi}_i: {\cal U}_i \times \bR \rightarrow
\bR^{n+1}$ by $\tilde{\chi}_i(\bx,t) = (\chi_i(\bx),t)$. Then $\{({\cal U}_i \times \bR, \tilde{\chi_i}):
i=1,...,N\}$ is a coordinate atlas for $\partial \Omega \times \bR \subset \bR^n$.
% From \cite{Tay:81}, Ch. II, Theorem 5.1, for $u \in
%C_0^{\infty}({\cal U}_i)$,
%\begin{equation}
 % \label{eqn:pullback}
 % (H(u \circ \chi_i))\circ \chi_i^{-1} = \tilde{H}_i u,
%\end{equation}
%with $\tilde{H}_i \in OPS^0({\cal U}_i)$ with principal symbol
%\begin{equation}
 % \sigma(\tilde{H}_i)_0(\bx,{\bf \xi}) =
  %\sigma(H)_0(\chi_i^{-1}(\bx),D\chi_i(\chi_i^{-1}(\bx))^T{\bf \xi})
%\end{equation}
%Also $\tilde{H}_i$ is essentially supported in $\chi_i^*\Gamma_1$

Every point $\bx \in \overline{\cal U}_i \cap \partial \Omega$ has a
neighborhoods $\bx \in {\cal U}\subset \overline{\cal U} \subset {\cal
  V}$ in which the
restrictions of $(\tilde{\chi}_i^{-1})^*\Gamma$ to $\chi_i({\cal V}) \times \bR$ are submanifolds of product conic
submanifolds $(\chi_i({\cal V}) \times \bR) \times \gamma_i
\subset T^*(\bR^{n+1})$, so that the pull-backs
$\tilde{\chi}_i^*((\chi_i({\cal V}) \times \bR) \times \gamma_i)
\supset \Gamma|_{{\cal V} \times \bR}$ 
satisfy condition 2 in Definition \ref{thm:definc} with $\delta$
replaced by $\delta/2$. 
Taking a finite subcover and intersections, identify the base open sets ${\cal U}$ with
the coordinate neigborhoods ${\cal U}_i$, with corresponding ${\cal
  V}_i \supset \overline{\cal U}_i$.
Choose a partition of unity $\{\phi_i: i=1,...,N\}$ subordinate to
$\{{\cal U}_i:i=1,...,N\}$, and for each $i=1,...N$, a cutoff function
$\psi_i\in C_0^{\infty}({\cal V}_i)$ that is $\equiv 1$ on ${\cal U}_i$.

The local product structure means that for each $i=1,...,N$, $({\bf \xi,}\omega) \in \gamma_{i}$,  condition 2 in
Definition \ref{thm:definc} is satisfied at $\tilde{\chi}_i^*(\bx,t,{\bf \xi},\omega)$ for all $(\bx,t) \in
(\chi_i({\cal V}_i) \times \bR)$  (with $\delta$ replaced by
$\delta/2$).
A standard local geometric asysmptotics construction implies the existence of a 
neighborhood of the closure of $(\partial \Omega \cap {\cal
  V}_i)\times \bR$,  in which for each $({\bf \xi,}\omega) \in
\gamma_{1,i}$, there exists a unique smooth solution of the eikonal
equation $\tau(\bx,t,{\bf \xi}/\omega)$ with $\tau_i(\bx,t {\bf \xi}/\omega) = t+\chi_i(\bx) \cdot {\bf
  \xi}/\omega$, for $\bx \in \partial \Omega \cap {\cal  V}_i$, $t \in
\bR$.

Choose $b \in C_0^{\infty}(\bR)$ so that $b(t)d(\bx,t)=d(\bx,t)$ for
$\bx \in \partial \Omega, t \in \bR$. Denote by $a_i$ the solution to order 2 of
the transport system  corresponding to $\tau_i$, with $a_i=b\psi_i$ on $(\partial \Omega \cap
{\cal  V}_i) \times \bR$. Because of
the smooth cut-off of $a_i$ at the lateral boundary of a ray tube, and
because of the uniformity of condition 2 in Definition
\ref{thm:definc} mentioned above,  the pair
$(\tilde{p}_i,\tilde{\bv}_i)$ given by
\begin{eqnarray}
  \label{eqn:pgo}
  \tilde{p}(\bx,t,{\bf \xi},\omega) &=& a_i(\bx,t,{\bf
                                        \xi}/\omega)e^{i\omega\tau(\bx,t,{\bf
                                        \xi}/\omega)},\\
  \label{eqn:vgo}
  \tilde{\bv}(\bx,t,{\bf \xi},\omega) & = & -\frac{1}{\rho(\bx)}\int_{-infty}^t dt \nabla \tilde{p}(\bx,t,{\bf \xi},\omega)
\end{eqnarray}
are well-defined and smooth in a neighborhood of $\partial \Omega
\times \bR$, and satisfy
\begin{eqnarray}
\label{eqn:aweloc}
\frac{1}{\kappa}\frac{\partial \tilde{p}_i}{\partial t} & = & - \nabla \cdot \tilde{\bv}_i +
\frac{1}{\omega^2}\tilde{r}_i, \nonumber \\
\rho\frac{\partial \tilde{\bv}_i}{\partial t} & = & - \nabla \tilde{p}_i
                                               \nonumber \\
\tilde{p}_i & =& 0 \mbox{ for }  t \ll 0,\nonumber\\ 
\tilde{bv}_i & = & 0 \mbox{ for } t \ll 0.
\end{eqnarray}
The remainder $\tilde{r}_i(\cdot,\cdot,{\bf \xi},\omega)$ is locally
$L^2$-bounded in a neighborhood of $\partial \Omega \times \bR$,
uniformly over $({\bf \xi},\omega) \in \gamma_{i}$.

Denote by $\tilde{\gamma}_i$ the complement of the n-dimensional unit
ball in $\gamma_i$. Let
\[
  \tilde{d}_i = (\phi_iHd) \circ \chi_i^{-1}
\]
The essential support of $H$ is strictly contained in
$\Gamma$, and $(\chi_i^{-1})^*\Gamma|_{{\cal U}_i} \subset
\chi_i({\cal U}_i) \times \gamma_i$, so the Fourier transform of
$\tilde{d}_i$ decreases rapidly outside of $\gamma_i$: that is,
\begin{equation}
  \label{eqn:smpart}
  \tilde{e}_i(\by,t) = \int_{\tilde{\gamma}_i^c} d\xi d\omega
  \hat{\tilde{d}_i}(\xi,\omega)e^{i(\by \cdot {\bf \xi} + \omega t)}
\end{equation}
is smooth. Recall that $a_i=b\psi_i=1$ on $\mbox{supp
}\phi_i Hd = \mbox{supp }\tilde{d}_i\circ \tilde{\chi}_i$, so for $\bx \in {\cal U}_i$
\[
(\tilde{d_i}- \tilde{e}_i) (\chi_i(\bx),t)  = \int_{\tilde{\gamma}_i} d\xi d\omega
\hat{\tilde{d}_i}(\xi,\omega)e^{i(\chi_i(\bx) \cdot {\bf \xi} + \omega  t)}
\]
\[
  = \int_{\tilde{\gamma}_i} d\xi d\omega
\hat{\tilde{d}_i}(\xi,\omega)(a_i(\bx,t,{\bf
  \xi},\omega)e^{i\tau_i(\bx,t,{\bf \xi}/\omega)} - \tilde{e}_i(\chi_i(\bx),t)
\]
\begin{equation}
  \label{eqn:prepsynth}
  = \int_{\tilde{\gamma}_i} d\xi d\omega \hat{\tilde{d}_i}({\bf
    \xi},\omega)\tilde{p}_i(\bx,t,{\bf \xi}, \omega) - \tilde{e}_i(\chi_i(\bx),t).
\end{equation}
Define 
\begin{equation}
  \label{eqn:psynth}
  \bar{p}_i(\bx,t) = \int_{\tilde{\gamma}_i} d\xi d\omega \hat{\tilde{d}_i}({\bf
    \xi},\omega)\tilde{p}_i(\bx,t,{\bf \xi}, \omega)
\end{equation}
for $\bx \in \partial \Omega \cap {\cal U}_i$. Extend the definition
of $\bar{p}_i$ to a neighborhood of $\partial \Omega$ in $\Omega$ by
imposing the condition \ref{eqn:psynth}, and
define $\bar{\bf{v}}_i$ in the same neighborhood by
\begin{equation}
  label{eqn:vsynth}
  \bar{\bf{v}}_i(\bx,t) = -\frac{1}{\rho(\bx)}\int_{-infty}^t dt \nabla \bar{p}_i(\bx,t)
\end{equation}
Then $(\bar{p}_i,{\bar{\bf v}}_i)$ solves the acoustic system with the right hand side
\begin{equation}
\label{eqn:rhs}
\bar{r}_i(\bx,t) = \int_{\tilde{\gamma}_i} d\xi d\omega
\frac{1}{\omega}^2\tilde{r}_i(\bx,t,{\bf \xi},\omega) -
\kappa(\bx)\frac{\partial}{\partial t}\tilde{e}_i(\chi_i(\bx),t) .
\end{equation}
On the boundary, for $\bx \in \partial \Omega$, the definition
\ref{eqn:psynth} implies that
\begin{equation}
  \label{eqn:locbc}
  \bar{p}_i(\bx,t) = \tilde{d}_i(\chi(\bx),t)
  =(\phi_iHd)(\bx,t)
\end{equation}
The sums
\begin{equation}
  \label{eqn:psynth2}
  p_2 = \sum_{i=1}^N\bar{p}_i, \bv_2 = \sum_{i=1}^N\bar{\bf v}_i
\end{equation}
solve the acoustic system with right-hand side
\begin{equation}
  \label{eqn:rsynth}
  r_2 = \sum_{i=1}^N\bar{r}_i
\end{equation}
in a neighborhood of the boundary $\partial \Omega \times \bR$, which
may be taken to be a collar neighborhood $\Omega_{\epsilon}$ for
sufficiently small $\epsilon$. Since $\{\phi_i: i=1,...,N\}$ is a
partition of unity, equation \ref{eqn:locbc} implies that
\begin{equation}
  \label{eqn:bc2}
  p_2|_{\partial \Omega \times \bR} = Hd.
\end{equation}

Since $\tilde{r}_i$ is uniformly $L^2$-bounded over $\gamma_i$ and
$\tilde{e}$ is smooth, $\partial_t^2\bar{r}_i$ is
locally square-integrable in $\Omega_{\epsilon}
\times \bR$. Denote by $\bu_3=(p_3,\bv_3) \in C^1(\bR,(L^2(\Omega))^{n+1})
\cap C^0(\bR,V)$ the strong causal solution of the system
$L\bu=(r_2,0)$. Since the restriction of
$p_3$ to $\partial \Omega \times \bR$ is well-defined and vanishes,
$\bu_4 = (p_4,\bv_4) = (p_2-p_3,\bv_2-\bv_3)$ is a causal weak
solution of $L\bu=0$ in $\Omega \times \bR$ with $p_4|_{\partial \Omega \times \bR}=Hd$.

 Let $\mu \in C_0^{\infty}(\bR^n)$ so that
$\mu=0$ in $\Omega_{\epsilon/2}$, $\mu=1$ in $\Omega - \Omega_{\epsilon}$.

\end{proof}

@@
